MA_Results

### 2024-01-24 Markdown/Manuscript/Child_RMDs/Clippings.Rmd
extracted ethics component

---
title: "Clippings"
output: html_document
date: "2023-04-18"
---

>>>>

---
title: "Clippings"
output: html_document
date: "2023-04-18"
---
<!-- Functions to implement model algorithms can be associated with a module via . -->

<!-- to provide tools for authoring CHEMs, supplying those CHEMs with data and using CHEMs to implement reproducible modelling analyses. -->

<!-- The ready4pack library depends on two other module authoring libraries. <!-- MOVED CONTENT -->

<!-- Each CHEM module authored with the ready4 framework will include both a data structure (specifying the required properties of data that can validly be supplied to a module) and a set of algorithms (specifying the operations that can be performed on data contained in a module instance). -->

<!-- development version -->

<!-- Talk about DARTH (https://doi.org/10.1007/s40273-019-00837-x )/ BCEA -->

### 2024-01-24 Markdown/Manuscript/Child_RMDs/MS_Aims.Rmd
close to submission update - except tables
missing tables
one table to go
extracted ethics component

(i)	propose core ethical responsibilities for CHEM developers and criteria for assessing ethical CHEM implementations;

(ii) describe a prototype software framework we have developed to support the ethical implementation of CHEMs; and

>>>>

(i)	describe a prototype software framework we have developed to support the ethical implementation of CHEMs; and
<!-- propose a set of ethical responsibilities for CHEM project teams and criteria for assessing ethical CHEM implementations; -->

<!-- (ii)  -->

### 2024-01-24 Markdown/Manuscript/Child_RMDs/MS_Cndl_Tabs_Not_In_Body.Rmd
extracted ethics component

```{r, child = '../Child_RMDs/MS_Tables_1.Rmd', echo=FALSE}
```
\newpage
```{r, child = '../Child_RMDs/MS_Tables_2.Rmd', echo=FALSE}
```
\newpage
```{r, child = '../Child_RMDs/MS_Tables_3.Rmd', echo=FALSE}
```
\newpage
```{r, child = '../Child_RMDs/MS_Tables_4.Rmd', echo=FALSE}
```

>>>>

<!-- # ```{r, child = '../Child_RMDs/MS_Tables_1.Rmd', echo=FALSE} -->
<!-- # ``` -->
<!-- # \newpage -->
<!-- # ```{r, child = '../Child_RMDs/MS_Tables_2.Rmd', echo=FALSE} -->
<!-- ``` -->

\newpage

```{r, child = '../Child_RMDs/MS_Tables_3.Rmd', echo=FALSE}
```

\newpage

```{r, child = '../Child_RMDs/MS_Tables_4.Rmd', echo=FALSE}
```

### 2024-01-24 Markdown/Manuscript/Child_RMDs/MS_Conclusion.Rmd
one table to go
extracted ethics component

---
title: "MS_Conclusion"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---
`r ifelse(is.na(params$X@conclusion_1L_chr),"",params$X@conclusion_1L_chr)` 
We have identified criteria that can be used to systematically assess the extent to which the computational implementation of health economic models adheres to transparency, reusability and updatability criteria. We have developed a prototype software framework that can support the implementation of CHEMs that largely meet these criteria. Our framework can be used as a prototype for developing future software frameworks to support ethical implementation of CHEMs.

>>>>

---
title: "MS_Conclusion"
author: "Author"
date: "`r Sys.Date()`"
output: html_document
---
`r ifelse(is.na(params$X@conclusion_1L_chr),"",params$X@conclusion_1L_chr)` 
<!-- We have identified criteria that can be used to systematically assess the extent to which the computational implementation of health economic models adheres to transparency, reusability and updatability criteria.  -->
We have developed a prototype software framework that can support the implementation of CHEMs that largely meet explicit criteria for CHEM transparency, reusability and updataility. Our framework can be used as a prototype for developing future software frameworks to support ethical implementation of CHEMs. Barriers to ethical CHEM practice arising from resource constraints should be addressed by research funders.

### 2024-01-24 Markdown/Manuscript/Child_RMDs/MS_Discussion.Rmd
extracted ethics component

<!-- Talk about DARTH (https://doi.org/10.1007/s40273-019-00837-x )/ BCEA -->
Ethical practice is a core expectation of health researchers and computational methods underpin most quantitative research, yet an understanding of what constitutes ethical public health modelling practice is underdeveloped [@10.3389/fpubh.2017.00068]. We have addressed this gap with our proposed responsibilities for model developers and funders, enabling CHEM attributes and CHEM implementation assessment criteria.

<!-- The less numerous and more focused assessment criteria we propose may potentially be more tractable to implement in reviews of models authored by third parties and as the basis for designing software frameworks to support ethical computational model implementation. -->

<!-- Two issues for funders - allow support for beneficial use, foster viable communities. -->

Currently, many if not most existing CHEMs are insufficiently transparent [@Jalali2021; @McManus2019; @Bermejo2017; @Ghabri2019], not reusable [@Feenstra2022; @Emerson2019], nor updatable [@Sampson2017; @10.3389/fpubh.2022.899874]. <!-- DELETION --> Addressing these issues will require changes to the tools, practices and funding underpinning health economic model development.

@@

Our software framework provides the outline of a technical solution to some of the challenges of open source CHEM development. However, better tools alone are insufficient and policy and institutional changes are also required. Reducing waste in research is a core responsibility of research funders [@chalmers2014increase] and funding the development of CHEMs that are not adequately understood, reused or updated is wasteful. Existing incentive structures for health economists generally do not promote facilitating peers to reuse their work and funding is rarely adequate to maintain and routinely update a model. Currently, it can take “an extraordinary amount of idealism” to commit to authoring and maintaining research software [@anzt2020environment]. Previously recommended strategies for more beneficial health economic research investments include support for harmonised ethical standards for model development [@10.3389/fpubh.2017.00068], methodological innovation to improve model transferability [@craig2018taking], networks of modellers working on common health conditions [@Sampson2019], and  centralized infrastructure such as open source model repositories [@Pouwels2022] and a standard platform for model implementations [@Ghabri2019]. Development of software frameworks to support ethical CHEM implementations could enable and enhance each of these strategies.

A future software framework for ethical CHEMs would ideally incorporate a base set of features useful to developers of computational models across all domains of public health and health economics, with the capability for community-led extensions that are tailored to the needs of modellers focused on specific health-conditions. Such approaches will likely be important to ensure a sufficient potential user-base to make CHEM software frameworks sustainable.

>>>>

<!-- Ethical practice is a core expectation of health researchers and computational methods underpin most quantitative research, yet an understanding of what constitutes ethical public health modelling practice is underdeveloped [@10.3389/fpubh.2017.00068]. We have addressed this gap with our proposed responsibilities for model developers and funders, enabling CHEM attributes and CHEM implementation assessment criteria.  -->

<!-- Currently, many if not most existing CHEMs are insufficiently transparent [@Jalali2021; @McManus2019; @Bermejo2017; @Ghabri2019], not reusable [@Feenstra2022; @Emerson2019], nor updatable [@Sampson2017; @10.3389/fpubh.2022.899874]. Addressing these issues will require changes to the tools, practices and funding underpinning health economic model development. -->

@@

Our software framework provides the outline of a technical solution to some of the challenges of open source CHEM development. However, better tools alone are insufficient and policy and institutional changes are also required. <!-- Reducing waste in research is a core responsibility of research funders [@chalmers2014increase] and funding the development of CHEMs that are not adequately understood, reused or updated is wasteful. Existing incentive structures for health economists generally do not promote facilitating peers to reuse their work and funding is rarely adequate to maintain and routinely update a model. Currently, it can take “an extraordinary amount of idealism” to commit to authoring and maintaining research software [@anzt2020environment]. Previously recommended strategies for more beneficial health economic research investments include support for harmonised ethical standards for model development [@10.3389/fpubh.2017.00068], methodological innovation to improve model transferability [@craig2018taking], networks of modellers working on common health conditions [@Sampson2019], and  centralized infrastructure such as open source model repositories [@Pouwels2022] and a standard platform for model implementations [@Ghabri2019]. Development of software frameworks to support ethical CHEM implementations could enable and enhance each of these strategies. --> A future software framework for ethical CHEMs would ideally incorporate a base set of features useful to developers of computational models across all domains of public health and health economics, with the capability for community-led extensions that are tailored to the needs of modellers focused on specific health-conditions. Such approaches will likely be important to ensure a sufficient potential user-base to make CHEM software frameworks sustainable.

### 2024-01-24 Markdown/Manuscript/Child_RMDs/MS_Imp_Example.Rmd
extracted ethics component

<!-- Add intro text -->
We are using the ready4 software framework to implement multiple transparent, reusable and updatable CHEMs in youth mental health (see Figure \@ref(fig:fig1)).  Some initial outputs from this work are publicly available as development version releases.

@@

We created a checklist (Table \@ref(tab:checktb)) that we used to subjectively assess these study outputs against TRU criteria. For each criterion, we provided a global assessment of whether it was met using the responses “yes”, “no” or “partial”.  We believe our utility mapping study CHEMs have satisfactorily met five of the six criteria (T1, T2, R1, R2 and U1) and have partially met one criterion (U2). The main shortcomings that we identified when applying the assessment criteria was that we have yet to adequately implement unit testing of the R libraries authored as part of this study.

>>>>

We are using the ready4 software framework to implement multiple transparent, reusable and updatable CHEMs in youth mental health (see Figure \@ref(fig:fig1)).  Some initial outputs from this work are publicly available as development version releases.

@@

We created a checklist (Table \@ref(tab:checktb)) that we used to subjectively assess these study outputs against <!-- Edit_start -->explicit criteria for assessing transparent, reusable and updatable CHEMs that we have previously developed (add REF) <!-- Edit_end -->. For each criterion, we provided a global assessment of whether it was met using the responses “yes”, “no” or “partial”.  We believe our utility mapping study CHEMs have satisfactorily met five of the six criteria (T1, T2, R1, R2 and U1) and have partially met one criterion (U2). The main shortcomings that we identified when applying the assessment criteria was that we have yet to adequately implement unit testing of the R libraries authored as part of this study.

### 2024-01-24 Markdown/Manuscript/Child_RMDs/MS_Imp_Toolkit.Rmd
extracted ethics component

To support the development of CHEMs that meet TRU assessment criteria, we have created a prototype software framework called ready4. The software framework aims to provide a toolkit for: (i) enabling modular implementation of CHEMs; (ii) authoring and documenting CHEM modules; (iii) managing the labelling and transfer of CHEM input and output data; and (iv) authoring reproducible analyses that apply CHEMs to compatible data.

To achieve these goals, we have implemented the software framework as R [@RCORE2022] code libraries that integrate with a number of online services and which are supported by a documentation website.
<!-- We have designed this framework to enable ethical computational implementations of health economic models that are both open-source (for transparency and reusability) and modular (for reusability and updatability). The framework aims to support the standardization and interoperability that are core enablers of implementing a modular and open source approaches.  -->

@@

<!-- Functions to implement model algorithms can be associated with a module via . -->

<!-- to provide tools for authoring CHEMs, supplying those CHEMs with data and using CHEMs to implement reproducible modelling analyses. -->

<!-- The ready4pack library depends on two other module authoring libraries. <!-- MOVED CONTENT -->

<!-- Each CHEM module authored with the ready4 framework will include both a data structure (specifying the required properties of data that can validly be supplied to a module) and a set of algorithms (specifying the operations that can be performed on data contained in a module instance). -->

<!-- development version -->

@@

##	Online services 
Framework libraries are designed to be used in conjunction with a number of third-party online services that we established and configured accounts with. 
<!-- To facilitate its application to our youth mental health project, we established and configured accounts with these required services (see Availability of data and materials). -->

>>>>

To support the development of CHEMs that are Transparent, Reusable and Updatable (TRU),<!-- assessment criteria,  --> we have created a prototype software framework called ready4. The software framework aims to provide a toolkit for: (i) enabling modular implementation of CHEMs; (ii) authoring and documenting CHEM modules; (iii) managing the labelling and transfer of CHEM input and output data; and (iv) authoring reproducible analyses that apply CHEMs to compatible data.

To achieve these goals, we have implemented the software framework as R [@RCORE2022] code libraries that integrate with a number of online services and which are supported by a documentation website.

@@

@@

##	Online services 
Framework libraries are designed to be used in conjunction with a number of third-party online services that we established and configured accounts with.

### 2024-01-24 Markdown/Manuscript/Child_RMDs/MS_Intro_Short.Rmd
one table to go
extracted ethics component

Health economics is a discipline concerned with problems that arise due to scarce resources, such as how to value health and healthcare, allocate healthcare budgets and configure health services [@wagstaff2012four]. In seeking to solve these problems, health economists typically use models which are simplified and selective representations of systems that are believed to influence human health. A health economic model should be capable of being described using words and figures (a conceptual model), equations (a mathematical model) and software (a computational model).  Health economic scientific manuscripts typically describe a model using its conceptual and mathematical representations but report results that have been generated by its computational representation (i.e., the format that allows computers to perform calculations). The conceptual, mathematical and computational representations of a model are assumed to be isomorphic. However, the ease of independently assessing the validity of this assumption depends in part on how a health economic model has been implemented computationally.
<!-- through execution of a computer program that applies the computational model to selected input data). -->
<!-- the execution of a computer program that applies a computational model to compatible data inputs (e.g., parameter values) and performs a sequence of numeric calculations. -->
<!-- the predictions reported in health economic studies are now almost always -->

The set of software files implementing a health economic model's structure, data and algorithms can be called a computational health economic model (CHEM). CHEMs can be developed using commercial modelling software or open-source programming languages such as Python [@python2009] or R [@RCORE2022]. An advantage of commercial modelling software is that users often require no or limited programming skills to develop and apply CHEMs. However, open-source software programming languages may facilitate the development of CHEMs that are more transparent, reusable and updatable [@incerti2019r; @Pouwels2022]. The decision about what type of software development platform to use must balance ethical considerations and resource constraints (see Table \@ref(tab:proscons)).

<!-- modelling team value judgments about priority CHEM features and the skills, resource and time constraints within which a CHEM project needs to be delivered -->

@@

Open-source CHEMs remain rare [@Feenstra2022; @Emerson2019]. The software used to develop CHEMs was identified as a barrier to open-source model implementations by participants in a 2020 survey of health economists [@Pouwels2022]. Other reported barriers relate to difficulties in generalising models, updating models and transferring model data, concerns about the level of public access and lack of interest from decision-makers, and legal, confidentiality and security considerations [@Pouwels2022]. 
<!-- Additional resources, skills and tools may be required to overcome a number of barriers to greater health economist adoption of open-source development. -->
<!-- A major consideration in open-source software development, is the selection of an appropriate software framework.  -->
<!-- on considerations such as model comprehensiveness, generalisability, reproducibility and transferability -->
<!-- Although there appears to be in principle support from many health economists for greater use of open source health economic models (OSHEMs)  [@Pouwels2022], actual implementations are rare [@Emerson2019; @Michalczyk2018; @Feenstra2022]. Barriers to adoption of OSHEMs include concerns about intellectual property, confidentiality, model misuse and the resources required to support open source implementations [@Pouwels2022; @Wu2019]. Adherence to good practice guidance is an essential requirement for healthcare modelling [@Erdemir2020],  -->

Better software frameworks could potentially help address some of the barriers to open-source CHEM development. A software framework is a shared common technology used by developers to collaboratively author software and which is not typically visible to software end-users [@myllarniemi2018development]. A software framework provides a foundation for developing multiple software applications with shared resources (e.g., code and data files), that can be modified to suit specific needs. Advantages of using software frameworks include facilitating code reuse and extension, promoting good programming practice and the capability to provide enhanced functionality and performance without additional effort by developers [@edwin2014software]. Software frameworks have widely been developed and implemented in data science, for example, PyTorch [@PyTorch2019] for machine learning using Python.

<!-- We are developing a model to explore multiple economic questions relating to the mental health of young people aged 12 to 25.   -->
<!-- When making choices about how to implement this health economic model computationally, we wished to facilitate ethical development and use. -->

A high-level software framework for implementing open-source CHEMs in R has previously been developed with the primary aim of improving model transparency [@alarid2019need]. However, to gain widespread adoption, a software framework for CHEMs may need to provide tools to help model developers meet a broader range of ethical practice requirements. There are multiple challenges to overcome. Firstly, the ethical responsibilities of public health computational modellers remain poorly understood <!-- and inadequately fulfilled  [@pliakos2021ethics; @thompson2022escape;]-->[@10.3389/fpubh.2017.00068] and the computational modelling landscape is changing fast [@calder2018computational]. Secondly, software frameworks can be challenging and time consuming to create and may become excessively complex over time [@edwin2014software]. Thirdly, software frameworks can be difficult to learn, often requiring model developers to undergo specialist training [@myllarniemi2018development].

<!-- a software framework is unlikely to be widely adopted unless the benefits to developers outweigh the costs associated with using it. -->
<!-- However, guidance on developing, selecting and using software frameworks for ethical CHEM implementations is scarce. -->
<!--  -->
<!-- Transparency has been proposed as one of the core ethical responsibilities of public health computational modellers [@10.3389/fpubh.2017.00068].  -->
<!-- Although computational models are widely used to inform health policy and system design [@dakin2015influence; @Erdemir2020; @LO2022100570; @Christene006827], -->

>>>>

<!-- Health economics is a discipline concerned with problems that arise due to scarce resources, such as how to value health and healthcare, allocate healthcare budgets and configure health services [@wagstaff2012four]. In seeking to solve these problems, health economists typically use models which are simplified and selective representations of systems that are believed to influence human health. A health economic model should be capable of being described using words and figures (a conceptual model), equations (a mathematical model) and software (a computational model).  Health economic scientific manuscripts typically describe a model using its conceptual and mathematical representations but report results that have been generated by its computational representation (i.e., the format that allows computers to perform calculations). The conceptual, mathematical and computational representations of a model are assumed to be isomorphic. However, the ease of independently assessing the validity of this assumption depends in part on how a health economic model has been implemented computationally. -->

<!-- The set of software files implementing a health economic model's structure, data and algorithms can be called a computational health economic model (CHEM). CHEMs can be developed using commercial modelling software or open-source programming languages such as Python [@python2009] or R [@RCORE2022]. An advantage of commercial modelling software is that users often require limited or no programming skills to develop and apply CHEMs. However, open-source software programming languages may facilitate the development of CHEMs that are more transparent, reusable and updatable [@incerti2019r; @Pouwels2022]. The decision about what type of software development platform to use must balance ethical considerations and resource constraints (see Table \@ref(tab:proscons)).  -->

@@

<!-- Open-source CHEMs remain rare [@Feenstra2022; @Emerson2019]. The software used to develop CHEMs was identified as a barrier to open-source model implementations by participants in a 2020 survey of health economists [@Pouwels2022]. Other reported barriers relate to difficulties in generalising models, updating models and transferring model data, concerns about the level of public access and lack of interest from decision-makers, and legal, confidentiality and security considerations [@Pouwels2022].  -->

<!-- Better software frameworks could potentially help address some of the barriers to open-source CHEM development.  -->
A software framework is a shared common technology used by developers to collaboratively author software and which is not typically visible to software end-users [@myllarniemi2018development]. A software framework provides a foundation for developing multiple software applications with shared resources (e.g., code and data files), that can be modified to suit specific needs. Advantages of using software frameworks include facilitating code reuse and extension, promoting good programming practice and the capability to provide enhanced functionality and performance without additional effort by developers [@edwin2014software]. Software frameworks have widely been developed and implemented in data science, for example, PyTorch [@PyTorch2019] for machine learning using Python.

A high-level software framework for implementing open-source <!-- Modification_start -->Computational Health Economic Models (CHEMs)<!-- Modification_end --> in R has previously been developed with the primary aim of improving model transparency [@alarid2019need]. However, to gain widespread adoption, a software framework for CHEMs may need to provide tools to help model developers meet a broader range of ethical practice requirements. There are multiple challenges to overcome. Firstly, <!-- the ethical responsibilities of public health computational modellers remain poorly understood [@10.3389/fpubh.2017.00068] and  -->the computational modelling landscape is changing fast [@calder2018computational]. Secondly, software frameworks can be challenging and time consuming to create and may become excessively complex over time [@edwin2014software]. Thirdly, software frameworks can be difficult to learn, often requiring model developers to undergo specialist training [@myllarniemi2018development].

### 2024-01-24 Markdown/Manuscript/Child_RMDs/MS_Tables.Rmd
extracted ethics component

```{r echo=FALSE}
if(params$X@tables_in_body_lgl){
  child_docs_ls <- list(one = "../Child_RMDs/MS_Tables_1.RMD",
                        two = "../Child_RMDs/MS_Tables_2.RMD",
                        three = "../Child_RMDs/MS_Tables_3.RMD")
}else{
    child_docs_ls <- list(one = NULL,
                      two = NULL)
}
```

>>>>

```{r echo=FALSE}
if(params$X@tables_in_body_lgl){
  child_docs_ls <- list(
    one = NULL,
    # "../Child_RMDs/MS_Tables_1.Rmd",
                        two = NULL,
    # "../Child_RMDs/MS_Tables_2.Rmd",
                        three = "../Child_RMDs/MS_Tables_3.Rmd",
                        four = "../Child_RMDs/MS_Tables_4.Rmd")
}else{
    child_docs_ls <- list(one = NULL,
                          two = NULL,
                          three = NULL,
                          four = NULL)
}
```
